Machine Learning guided workflow for Ribosome Binding Site engineering

### 2021-05-28 paper/archive_supp.tex
re-arange supp; plots

\textbf{Supplementary}

\section{Extended Machine Learning Discussion}

\subsection{Reproducible Plots}
In this section, we put reproducible plots and will update the plots when we get full results.

\begin{figure}[h]
    \centering
    \includegraphics[scale = 0.4]{plots/Supplementary/swarmplots.pdf}
    \caption{Swarmplots for different groups, with raw (left) and normalised (right) TIR labels (averaged over 6 replicates). 
    Group names represent: Consensus (consensus sequence tested in different round); BPS-NC (bps noncore); BPS-C (bps core); UNI (uniformly random); PPM (position-based probability matrix); Bandit-0 (bandit design for round 0); Bandit-1 (bandit design for round 1).}
    \label{fig:swarmplots.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale = 0.3]{plots/Supplementary/Data_Similarity_and_Prediction_Sorted_by_TIR_Label.pdf}
    \includegraphics[scale = 0.3]{plots/Supplementary/Data_Similarity_and_Prediction_Sorted_by_RBS_Similarity.pdf}
    \caption{Kernel Heatmap and Predictions. Sequences are grouped as Consensus (1-2), BPS-C (3-20), BPS-NC (21-61), UNI (62-90), PPM (91-118), Bandit-0 (119-177), Bandit-1 (178-265). Inside of each group, sequences are clustered and sorted in terms of TIR labels (left) or RBS similarity (right). The first row shows the similarity measured by weighted degree kernel with shift, the second shows the predicted mean and uncertainty (1.95 standard deviation).}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale = 0.25]{plots/Supplementary/clustering.png}
    \caption{TSNE of RBS sequences in design space with clustering. The distance is calculated based on the weighted degree kernel on RBS sequences. Colours indicate different groups, shapes indicate different clusters.}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.4]{plots/Supplementary/core_vs_noncore.pdf}
    \caption{\textbf{Comparison of base change impact on TIR in core versus non-core region.} The core region is highligthed in light green and the lines are rolling averages for each base. The top dotted line shows the TIR for the benchmark sequence, where dots represent a change at a given position to a given base, which is colour coded.}
    \label{fig:core_vs_noncore}
\end{figure}

@@

\section{Supplementary Figures}

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[scale=0.25]{plots/Supplementary/Structure_GGGGGG.pdf}
         \caption{$y=x$}
         \label{fig:GGGGGG}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[scale=0.25]{plots/Supplementary/Structure_GGGGGC.pdf}
         \caption{$y=3sinx$}
         \label{fig:GGGGGC}
     \end{subfigure}
     \vskip\baselineskip
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[scale=0.25]{plots/Supplementary/Structure_CGGGAC.pdf}
         \caption{$y=5/x$}
         \label{fig:CGGGAC}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[scale=0.25]{plots/Supplementary/Structure_CACGTT.pdf}
         \caption{$y=5/x$}
         \label{fig:CACGTT}
     \end{subfigure}
        \caption{\textbf{Folding predictionds for four different RBS.} Here we show the predicted, energetically most favourable structures for four of our RBS with the following cores: GGGGGG, GGGGGC, CGGGAC, CACGTT and their immediate upstream and downstream background sequence. There is no discerbile and consistent difference between strong ones (the first two) and weak ones (the last two).}
        \label{fig:structures}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{plots/Supplementary/All_logo.pdf}
    \caption{\textbf{Sequence logo for all tested sequences.} }
    \label{fig: All_logo}
\end{figure}

\begin{figure}[!ht]
    \centering
     \includegraphics[scale=0.4]{plots/Supplementary/proj_scatter_abc1_FF.pdf}
    \caption{Exploration v.s. Exploitation.}
\end{figure}

\begin{figure}[!ht]
    \centering
     \includegraphics[scale=0.4]{paper/plots/Supplementary/SDhist.pdf}
    \caption{Histogram of standard deviations.}
\end{figure}

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \caption{}
         \includegraphics[scale=0.4]{plots/Supplementary/SD.pdf}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \caption{}
         \includegraphics[scale=0.6]{plots/Supplementary/SDvsTIR.pdf}
     \end{subfigure}
     \caption{\textbf{Distribution of standard deviation of tested samples} A) Joint plot showing the distribution of relative error calculated as the average TIR from 6 replicates divided by the given samples standard deviation (coefficient of variation). B) Scatter plot showing the standard deviation for each RBS plotted against its averaged TIR. Standard deviation and average TIR is given in terms of raw number.}
     \label{fig:variation of biological replicates}
\end{figure}

\input{mixture_absGau}

>>>>

@@

\begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.4]{plots/Supplementary/core_vs_noncore.pdf}
    \caption{\textbf{Comparison of base change impact on TIR in core versus non-core region.} The core region is highligthed in light green and the lines are rolling averages for each base. The top dotted line shows the TIR for the benchmark sequence, where dots represent a change at a given position to a given base, which is colour coded.}
    \label{fig:core_vs_noncore}
\end{figure}

@@

\subsection{Mixture of Absolute Gaussian}

The raw TIR scores in each round are positive and  have two modes. 
We assume they are mixture of two absolute Gaussian distributions. 
If we can fit the mixture of absolute Gaussian distributions, we can answer the following questions:
\begin{itemize}
    \item is the left peak decreasing in height with each round?
	\item is the left peak going left?
	\item Is the right peak increasing in height with each round?
	\item is the right peak going right?
\end{itemize}

The density of absolute Gaussian (also called \textit{folded normal distribution}) is 
\begin{align}
    f_{Y}\left(x ; \mu, \sigma^{2}\right)=\frac{1}{\sqrt{2 \pi \sigma^{2}}} e^{-\frac{(z-\mu)^{2}}{2 \sigma^{2}}}+\frac{1}{\sqrt{2 \pi \sigma^{2}}} e^{-\frac{(x+\mu)^{2}}{2 \sigma^{2}}}
\end{align}
for $x \geq 0$ and $0$ everywhere else. 
The mean and variance of the absolute Gaussian distribution is 
\begin{align}
    \mu_{Y} &=\sigma \sqrt{\frac{2}{\pi}} \exp \left(\frac{-\mu^{2}}{2 \sigma^{2}}\right)+\mu \operatorname{erf}\left(\frac{\mu}{\sqrt{2 \sigma^{2}}}\right);\\
    \sigma_{Y}^{2} &=\mu^{2}+\sigma^{2}-\mu_{Y}^{2},
\end{align}
where $\operatorname{erf}$ is the error function 
$\operatorname{erf} z=\frac{2}{\sqrt{\pi}} \int_{0}^{z} e^{-t^{2}} d t$.

We consider the model of the mixture of two absolute Gaussian distributions as

\begin{figure}[!ht]
    \centering
    \includegraphics[scale = 0.25]{plots/Supplementary/clustering.png}
    \caption{TSNE of RBS sequences in design space with clustering. The distance is calculated based on the weighted degree kernel on RBS sequences. Colours indicate different groups, shapes indicate different clusters.}
\end{figure}

### 2021-05-28 paper/methods.tex
Updates from Overleaf
re-arange supp; plots

%  For recommendations, we consider the \textit{Upper Confidence Bound (UCB)} type algorithms. 
%  As one popular type of the bandit algorithms \cite{lattimore2018bandit}, the UCB type of algorithms are based on the \textit{optimism in the face of uncertainty},
 %provide various approaches to sequential design where an agent adaptively chooses one or more options among several actions based on certain policies. In our work we used the Upper Confidence Bound version of that algorithm, which is based on the \textit{optimism in the face of uncertainty}. The UCB algorithm, as the name suggests,
%  which basically select RBS sequences with the maximum upper confidence bound constructed by the sum of the predicted mean and $n$ standard deviation ($n > 0$), i.e. $\operatorname{argmax}_{\mathbf{x}_i \in \mathcal{D}} \left( \mu_t(\mathbf{x}_i) + \beta_t \sigma_t(\mathbf{x}_i)\right)$,
%     where $\beta_t$ is a hyperparameter balancing the exploitation and exploration, 
%     $\mu_t(\mathbf{x}_i), \sigma_t(\mathbf{x}_i)$ are the predicted mean and standard deviation at round $t$ for the sequence $\mathbf{x}_i$.

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=.4]{plots/Recommend_2_Data_Point_Top_2_UCB.pdf}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/GP-BUCB_Recommend_1st_Data_Point.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/GP-BUCB_Recommend_2nd_Data_Point.pdf}
    \end{subfigure}
    \caption{Batch Recommendation Illustration. We use the batch size of 2, with 5 initial observations. The design space is 24 uniformly distributed points in the range [-2,10], i.e. {-2, -1.5, -1, ..., 9.5, 10}.
    The confidence interval are shown with predicted mean of $\pm$ 1.96 standard deviation.
    (A) Top UCB recommendations. The recommendations are 2 data points with top UCB scores, constructed with GP predictions. 
    (B)(C) Batch UCB recommendations. (B) shows the first recommended sequence, (C) shows the new predicted confidence interval and the second recommendation based on that.}
    \label{fig:batch rec}
\end{figure}

>>>>

%  For recommendations, we consider the \textit{Upper Confidence Bound (UCB)} type algorithms. 
%  As one popular type of the bandit algorithms \cite{lattimore2020bandit}, the UCB type of algorithms are based on the \textit{optimism in the face of uncertainty},
 %provide various approaches to sequential design where an agent adaptively chooses one or more options among several actions based on certain policies. In our work we used the Upper Confidence Bound version of that algorithm, which is based on the \textit{optimism in the face of uncertainty}. The UCB algorithm, as the name suggests,
%  which basically select RBS sequences with the maximum upper confidence bound constructed by the sum of the predicted mean and $n$ standard deviation ($n > 0$), i.e. $\operatorname{argmax}_{\mathbf{x}_i \in \mathcal{D}} \left( \mu_t(\mathbf{x}_i) + \beta_t \sigma_t(\mathbf{x}_i)\right)$,
%     where $\beta_t$ is a hyperparameter balancing the exploitation and exploration, 
%     $\mu_t(\mathbf{x}_i), \sigma_t(\mathbf{x}_i)$ are the predicted mean and standard deviation at round $t$ for the sequence $\mathbf{x}_i$.

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=.4]{plots/Main_Paper/Recommend_2_Data_Point_Top_2_UCB.pdf}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Main_Paper/GP-BUCB_Recommend_1st_Data_Point.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Main_Paper/GP-BUCB_Recommend_2nd_Data_Point.pdf}
    \end{subfigure}
    \caption{Batch Recommendation Illustration. We use the batch size of 2, with 5 initial observations. The design space is 24 uniformly distributed points in the range [-2,10], i.e. {-2, -1.5, -1, ..., 9.5, 10}.
    The confidence interval are shown with predicted mean of $\pm$ 1.96 standard deviation.
    (A) Top UCB recommendations. The recommendations are 2 data points with top UCB scores, constructed with GP predictions. 
    (B)(C) Batch UCB recommendations. (B) shows the first recommended sequence, (C) shows the new predicted confidence interval and the second recommendation based on that.}
    \label{fig:batch rec}
\end{figure}
